\textcolor{Navy}{\textsc{Convexity adjustments with a bit of Malliavin}}

### 2023-03-18 Latex/paper-resum.tex
# last version of the paper where we have fixed some notation typos.

In order to introduce a general idea of the method. Let define a $Z_t = f(x_t)$. Now, we suppose that $Z_t$ is martingale under the measure $\mathbb{Q}_1$ and that we want to compute $\mathbb{E}^{\mathbb{Q}_2}\left(Z_T \right)$, where $\mathbb{Q}_2$ is a measure such that $dW^{\mathbb{Q}_1}_t = dW^{\mathbb{Q}_2}_t +\lambda_t dt$. Then, if we use the Clark-Ocone representation, we have that
\begin{equation*}
f(x_t) = \mathbb{E}^{\mathbb{Q}_1}\left(f(x_t)\right) + \int_{0}^{t} \mathbb{E}^{\mathbb{Q}_1}_s\left( f^{\prime}(x_t) D_s x_t  \right) dW^{\mathbb{Q}_1}_s
\end{equation*}
Now, if we take $\mathbb{E}^{\mathbb{Q}_2}\left( \cdot \right)$ in the previous expression and we use the Girsanov's theorem, we get that
\begin{equation}\label{general_convexity}
\mathbb{E}^{\mathbb{Q}_2}\left( f(x_t) \right) = f(x_0) + \mathbb{E}^{\mathbb{Q}_1} \left(\int_{0}^{t}  \mathbb{E}^{\mathbb{Q}_1}_s\left( f^{\prime}(x_t) D_s x_t  \right) \lambda_s ds \right). 
\end{equation}
The second term, is the convexity adjustment to due to change of measure from $\mathbb{Q}_1$ to $\mathbb{Q}_2$. The differents choice of $f$, $\mathbb{Q}_1$ and $\mathbb{Q}_2$ will allow us to get an approximation of the convexity adjustment for the different case of interest. 
\subsection{FRAs Vs futures}
 The cash flows in FRAs and futures are computed under different measures. Consequently, we need to adjust the futures price quote to transform them into FRAs price quotes. As usual, we will define the forward rate at time $t_0$ between $t_1$ and $t_2$ under the forward curve $E$ as:
\begin{equation}\label{forward_rate}
L_{E}(t_0, t_1, t_2) = \frac{1}{\delta_{t_1,t_2}}\left(\frac{P_{E}(t_0,t_1)}{P_{E}(t_0,t_2)} - 1 \right)
\end{equation} 
where $P_{E}(t,T)$ is the discount factor for the curve $E$ from $t$ to $T$, and $\delta_{t_1,t_2}$ is the year fraction between $t_1$ and $t_2$.

@@

\begin{theorem}\label{Th_CA_OIS}[Convexity Adjustment approximation for OIS Futures]
Given the Cheyette model in \eqref{short_rate_cheyette}, the hypothesis \ref{boundedness_volatility} and \ref{boundedness_reversion}, and considering the approximations in \eqref{approximation_y_t} and \eqref{approximation_x_t_a}. Then, the convexity adjustment approximation for OIS futures is 
\begin{equation}\label{convexity_ois_future}
\mathbb{E}_t^{\mathbb{Q}}\left(R(t_0,t_1)\right) \approx \frac{\exp\left(\mathbb{E}_t^{\mathbb{Q}}\left(I(t_0,t_1)\right)\right)\exp\left(-\int_{t}^{t_1}\frac{\Gamma^{2}(s,t_0,t_1)}{2} ds\right) - 1}{\delta_{t_0,t_1}}
\end{equation}
and for the average OIS future is
\begin{equation}\label{convexity_avg_ois_future}
\mathbb{E}_t^{\mathbb{Q}}\left(R_{avg}(t_0,t_1)\right) = \frac{\mathbb{E}_t^{\mathbb{Q}}\left(I(t_0,t_1)\right) }{\delta_{t_0,t_1}} \approx \frac{\log\left(1+\delta_{t_0,t_1}  \mathbb{E}_t^{\mathbb{Q}}\left(R(t_0,t_1)\right) \right)}{\delta_{t_0,t_1}} - \frac{\int_{0}^{t_1}  \Gamma^{2}(s,t_0,t_1) ds}{2\delta_{t_0,t_1}}
\end{equation}
\end{theorem}
with  $\Gamma(s,t_0,t_1)= \eta(s,x_0,y_0)\int_{\max(s, t_{0})}^{t_1} \exp\left( -\int_{s}^{u} k_{s^{\prime}} ds^{\prime}\right)du$
\begin{proof}
See appendix \ref{Proof_CA_OIS_futures}.
\end{proof}

@@

The next figure, show accuracy of (\ref{convexity_ois_future}) and (\ref{convexity_avg_ois_future}) for the Hull-White model case when the mean reversion is $k=0.003$ and volatility $\sigma=0.01$. 
\begin{figure}[H]
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[scale=0.2]{Figures/convexity_ois.jpg}
		\caption{Convexity Mc Vs Convexity Malliavin}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[scale=0.2]{Figures/convexity_avg_ois.jpg}
		\caption{Convexity Avg Mc Vs Convexity Avg Malliavin}
\end{subfigure}
\caption{Convexity adjustment for compounding and average OIS}
\end{figure} 
\end{example}

@@

\begin{theorem}\label{Th_CA_CMS}[Convexity Adjustment approximation for CMS]
Given the Cheyette model in \eqref{short_rate_cheyette}, the hypothesis \ref{boundedness_volatility} and \ref{boundedness_reversion}, and considering the approximations in \eqref{approximation_y_t} and \eqref{approximation_x_t_a}. Given
\begin{equation}
M(t,t_p)= \frac{P_{ois}(t,t_p)}{01(t,t_a,t_p)}
\end{equation}
and 
\begin{align}\label{cms_first_order_convexity}
\mathbb{E}^{t_p}\left(S_{a,b}(t_a)\right) &\approx  S^{ois}_{a,b}(0) + \frac{\partial_x S_{a,b}(t_a,\bar{x}_0(t_a), \bar{y}_0) \partial_x M(t_a,t_p,\bar{x}_0(t_a), \bar{y}_0)} \nonumber \\
&\cdot \int_{0}^{t_a}  \beta^2(s,t_a,\bar{x}_0(t_a),\bar{y}_s) \exp\left(-2\int_{s}^{t_a}\beta(u,t_a,\bar{x}_{0}(t_a),\bar{y}_u) \mu(u,t_a,t_b) du \right)ds 
\end{align}
with  $\beta(u,t_a,x,y) = \exp\left(-\int_{u}^{t_a}k_u du\right)\eta(u,x,y)$. Then, the convexity adjustment approximation for CMS is
\begin{equation}
CA_{CMS}(t_p) = \mathbb{E}^{t_p}\left(S_{a,b}(t_a)\right) - S_{a,b}(0).
\end{equation} 
\end{theorem}
\begin{proof}
See appendix \ref{Proof_CA_CMS}.
\end{proof}

@@

\bar{DM}(s,t_a) &= \mathbb{E}^{t_p}_s\left(\exp\left(-\int_{s}^{t_a} \exp\left(-\int_{u}^{t_a}k_u du\right)\partial_x \eta(u, \bar{x}_0(t_a), \bar{y}_{u}) \bar{\nu}(u,t_p, \bar{x}_0(t_a)) \right. \right. \\
&+ \left. \left. \eta(u,\bar{x}_0(t_a), \bar{y}_{u})) \partial_x \bar{\nu}(t,t_p, \bar{x}_0(t_a)) du \right)\right) 

@@

\mathbb{E}_s(D_s \bar{x}_{t_a}) &\approx \eta(s,\bar{x}_s,\bar{y}_s) \exp\left(-\int_{s}^{t_a}k_u du \right)\\ &\cdot \mathbb{E}^{t_p}_s\left(\exp\left(-\int_{s}^{t_a} \exp\left(-\int_{u}^{t_a}k_u du\right)\partial_x (\eta(u, \bar{x}_0(t_a), \bar{y}_{u})) \bar{\nu}(u,t_p, \bar{x}_0(t_a)) \right. \right. \\
&+ \left. \left. \eta(u, \bar{x}_0(t_a), \bar{y}_{u})) \partial_x  \bar{\nu}(t,t_p, \bar{x}_0(t_a)) du \right)\right) 

@@

\subsection{Proof Theorem \ref{Th_CA_FRAsinArrears}}\label{Proof_CA_FRAsinArrears}
$L_{E}(t,t_1,t_2)$ is martingale under the measure $\mathbb{Q}^{t_2}$, therefore the expected value of \eqref{FRAinArrear} is taken with respect to the wrong martingale. To calculate the convexity adjustment we use the Clark-Ocone to get a representation for $L_{E}(t_1,t_1,t_2)$ i.e
\begin{equation}\label{general_convexity_fras}
L_{E}(t_1,t_1,t_2) = \mathbb{E}^{t_2}\left(L_{E}(t_1,t_1,t_2) \right) + \int_{0}^{t_1} \mathbb{E}^{t_2}\left(D_s L_{E}(t_1,t_1,t_2) \right) dW^{\mathbb{Q}^{t_2}}_{s}
\end{equation}
Under the HJM dynamics, we have the relation
\begin{equation*}
dW^{\mathbb{Q}^{t_2}}_s = dW^{\mathbb{Q}^{t_1}}_s + (\nu(s,t_2) - \nu(s,t_1)) ds. 
\end{equation*}
Taking $\mathbb{E}^{^{\mathbb{Q}^{t_1}}}(\cdot)$, we get that
\begin{align*}
\mathbb{E}^{t_1}\left(L_{E}(t_1,t_1,t_2) \right) &= L_{E}(0,t_1,t_2) + \mathbb{E}^{t_1}\left( \int_{0}^{t_1} \mathbb{E}_s^{t_2}\left(D_s L_{E}(t_1,t_1,t_2) \right) dW^{t_2}_{s}\right)\\
&= L_{E}(0,t_1,t_2) + \mathbb{E}^{t_1}\left( \int_{0}^{t_1} \mathbb{E}_s^{t_2}\left(D_s L_{E}(t_1,t_1,t_2) \right) (\nu(s,t_2)-\nu(s,t_1)) ds \right).
\end{align*}
Now from (\ref{approximation_D_s_x_t}) we have that 
\begin{eqnarray*}
D_s L(t_1,t_1,t_2) &=& \frac{G(t_1,t_2)}{\delta_{t_1,t_2}P_{E}(t_1,t_2)} D_s x_{t_1}\\
&\approx& \frac{G(t_1,t_2)}{\delta_{t_1,t_2}P_{E}(0,t_1,t_2)} \beta(s,t_1, \bar{x}_0(t_1), \bar{y}_s)\bar{DM}(s,t_1)
\end{eqnarray*}
If we use the above approximation, we get
\begin{equation*}
CA(t_0,t_1) = \frac{G(t_1,t_2)}{\delta_{t_1,t_2}P_{E}(0,t_1,t_2)} \beta(s,t_1, \bar{x}_0(t_1), \bar{y}_s) \int_{0}^{t_1} \bar{DM}(s,t_1) (\bar{\nu}(s,t_2,\bar{x}_{0}(t_1))-\bar{\nu}(s,t_1,\bar{x}_{0}(t_1))) ds
\end{equation*}
with $\bar{\nu}(t,t_p, \bar{x}_0(t_a))= \int_{t}^{t_p} \eta(s,\bar{x}_0(t_a),\bar{y}_s) ds$.

>>>>

In order to introduce a general idea of the method. Let define a $Z_t = f(x_t)$. Now, we suppose that $Z_t$ is martingale under the measure $\mathbb{Q}_1$ and that we want to compute $\mathbb{E}^{\mathbb{Q}_2}\left(Z_T \right)$, where $\mathbb{Q}_2$ is a measure such that $dW^{\mathbb{Q}_1}_t = dW^{\mathbb{Q}_2}_t +\lambda_t dt$. Then, if we use the Clark-Ocone representation, we have that
\begin{equation*}
f(x_t) = \mathbb{E}^{\mathbb{Q}_1}\left(f(x_t)\right) + \int_{0}^{t} \mathbb{E}^{\mathbb{Q}_1}_s\left( f^{\prime}(x_t) D_s x_t  \right) dW^{\mathbb{Q}_1}_s
\end{equation*}
Now, if we take $\mathbb{E}^{\mathbb{Q}_2}\left( \cdot \right)$ in the previous expression and we use the Girsanov's theorem, we get that
\begin{equation}\label{general_convexity}
\mathbb{E}^{\mathbb{Q}_2}\left( f(x_t) \right) = f(x_0) + \mathbb{E}^{\mathbb{Q}_1} \left(\int_{0}^{t}  \mathbb{E}^{\mathbb{Q}_1}_s\left( f^{\prime}(x_t) D_s x_t  \right) \lambda_s ds \right). 
\end{equation}
The second term, is the convexity adjustment to due to change of measure from $\mathbb{Q}_1$ to $\mathbb{Q}_2$. The differents choice of $f$, $\mathbb{Q}_1$ and $\mathbb{Q}_2$ will allow us to get an approximation of the convexity adjustment for the different case of interest. 
\subsection{FRAs Vs futures}
 The cash flows in FRAs and futures are computed under different measures. Consequently, we need to adjust the futures price quote to transform them into FRAs price quotes. As usual, we will define the forward rate at time $t_0$ between $t_1$ and $t_2$ under the forward curve $E$ as:
\begin{equation}\label{forward_rate}
L_{E}(t_0, t_1, t_2) = \frac{1}{\delta_{t_1,t_2}}\left(\frac{P_{E}(t_0,t_1)}{P_{E}(t_0,t_2)} - 1 \right)
\end{equation} 
where $P_{E}(t,T)$ is the discount factor for the curve $E$ from $t$ to $T$, and $\delta_{t_1,t_2}$ is the year fraction between $t_1$ and $t_2$. we must note, that $L_{E}(t_0, t_1, t_2)$ is a martingale under the forward measure $\mathbb{Q}^{t_2}$ associated with the numeraire $P_{ois}(t,t_2)$.

@@

\begin{theorem}\label{Th_CA_OIS}[Convexity Adjustment approximation for OIS Futures]
Given the Cheyette model in \eqref{short_rate_cheyette}, the hypothesis \ref{boundedness_volatility} and \ref{boundedness_reversion}, and considering the approximations in \eqref{approximation_y_t} and \eqref{approximation_x_t_a}. Then, the convexity adjustment approximation for OIS futures is 
\begin{equation}\label{convexity_ois_future}
\bar{R}(t_0,t_1) \approx \frac{\exp\left(\mathbb{E}_t^{\mathbb{Q}}\left(I(t_0,t_1)\right)\right)\exp\left(-\int_{t}^{t_1}\frac{\Gamma^{2}(s,t_0,t_1)}{2} ds\right) - 1}{\delta_{t_0,t_1}}
\end{equation}
and for the average OIS future is
\begin{equation}\label{convexity_avg_ois_future}
\bar{R}_{avg}(t_0,t_1) = \frac{\mathbb{E}_t^{\mathbb{Q}}\left(I(t_0,t_1)\right) }{\delta_{t_0,t_1}} \approx \frac{\log\left(1+\delta_{t_0,t_1}  \mathbb{E}_t^{\mathbb{Q}}\left(R(t_0,t_1)\right) \right)}{\delta_{t_0,t_1}} - \frac{\int_{0}^{t_1}  \Gamma^{2}(s,t_0,t_1) ds}{2\delta_{t_0,t_1}}
\end{equation}
\end{theorem}
with  $\Gamma(s,t_0,t_1)= \eta(s,x_0,y_0)\int_{\max(s, t_{0})}^{t_1} \exp\left( -\int_{s}^{u} k_{s^{\prime}} ds^{\prime}\right)du$
\begin{proof}
See appendix \ref{Proof_CA_OIS_futures}.
\end{proof}

@@

The next figure, show accuracy of (\ref{convexity_ois_future}) and (\ref{convexity_avg_ois_future}). The parameters that we have used to run the Monte Carlo have been $k=0.003$ and volatility $\sigma=0.01$ and flat curve $r=0.01$. 
\begin{figure}[H]
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[scale=0.2]{Figures/convexity_ois.jpg}
		\caption{Convexity Mc Vs Convexity Malliavin}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[scale=0.2]{Figures/convexity_avg_ois.jpg}
		\caption{Convexity Avg Mc Vs Convexity Avg Malliavin}
\end{subfigure}
\caption{Convexity adjustment for compounding and average OIS}
\end{figure} 
\end{example}

@@

\begin{theorem}\label{Th_CA_CMS}[Convexity Adjustment approximation for CMS]
Given the Cheyette model in \eqref{short_rate_cheyette}, the hypothesis \ref{boundedness_volatility} and \ref{boundedness_reversion}, and considering the approximations in \eqref{approximation_y_t} and \eqref{approximation_x_t_a}. Given
\begin{equation}
M(t,t_p)= \frac{P_{ois}(t,t_p)}{01(t,t_a,t_p)}
\end{equation}
Then, we have the temporal convexity adjusment for a CMS rate is approximated by
\begin{align} \label{cms_first_order_convexity}
CA_{CMS}(t_p) &\approx \frac{\partial_x S_{a,b}(t_a,\bar{x}_0(t_a), \bar{y}_0) \partial_x M(t_a,t_p,\bar{x}_0(t_a), \bar{y}_0)}{M(0,t_p)} \nonumber \\
&\cdot \int_{0}^{t_a}  \beta^2(s,t_a,\bar{x}_0(t_a),\bar{y}_s) \exp\left(-2\int_{s}^{t_a}\beta(u,t_a,\bar{x}_{0}(t_a),\bar{y}_u) \mu(u,t_a,t_b) du \right)ds 
\end{align}
with  $\beta(u,t_a,x,y) = \exp\left(-\int_{u}^{t_a}k_u du\right)\eta(u,x,y)$.
\end{theorem}
\begin{proof}
See appendix \ref{Proof_CA_CMS}.
\end{proof}

@@

\bar{DM}(s,t_a) &= \mathbb{E}^{t_p}_s\left(\exp\left(-\int_{s}^{t_a} \exp\left(-\int_{u}^{t_a}k_u du\right)\left(\partial_x \eta(u, \bar{x}_0(t_a), \bar{y}_{u}) \bar{\nu}(u,t_p, \bar{x}_0(t_a)) \right. \right. \right. \\
&+ \left. \left. \left. \eta(u,\bar{x}_0(t_a), \bar{y}_{u})) \partial_x \bar{\nu}(t,t_p, \bar{x}_0(t_a)) du \right) \right)\right) 

@@

\mathbb{E}_s(D_s \bar{x}_{t_a}) &\approx \eta(s,\bar{x}_s,\bar{y}_s) \exp\left(-\int_{s}^{t_a}k_u du \right)\\ &\cdot \mathbb{E}^{t_p}_s\left(\exp\left(-\int_{s}^{t_a} \exp\left(-\int_{u}^{t_a}k_u du\right)\left(\partial_x (\eta(u, \bar{x}_0(t_a), \bar{y}_{u})) \bar{\nu}(u,t_p, \bar{x}_0(t_a)) \right. \right. \right. \\
&+ \left. \left. \left. \eta(u, \bar{x}_0(t_a), \bar{y}_{u})) \partial_x  \bar{\nu}(t,t_p, \bar{x}_0(t_a)) du \right)\right)\right) 

@@

\subsection{Proof Theorem \ref{Th_CA_FRAsinArrears}}\label{Proof_CA_FRAsinArrears}
$L_{E}(t,t_1,t_2)$ is martingale under the measure $\mathbb{Q}^{t_2}$, therefore the expected value of \eqref{FRAinArrear} is taken with respect to the wrong martingale. To calculate the convexity adjustment we use the Clark-Ocone to get a representation for $L_{E}(t_1,t_1,t_2)$ i.e
\begin{equation}\label{general_convexity_fras}
L_{E}(t_1,t_1,t_2) = \mathbb{E}^{t_2}\left(L_{E}(t_1,t_1,t_2) \right) + \int_{0}^{t_1} \mathbb{E}^{t_2}\left(D_s L_{E}(t_1,t_1,t_2) \right) dW^{\mathbb{Q}^{t_2}}_{s}
\end{equation}
Under the HJM dynamics, we have the relation
\begin{equation*}
dW^{\mathbb{Q}^{t_2}}_s = dW^{\mathbb{Q}^{t_1}}_s + (\nu(s,t_2) - \nu(s,t_1)) ds. 
\end{equation*}
Taking $\mathbb{E}^{^{t_1}}(\cdot)$, we get that
\begin{align*}
\mathbb{E}^{t_1}\left(L_{E}(t_1,t_1,t_2) \right) &= L_{E}(0,t_1,t_2) + \mathbb{E}^{t_1}\left( \int_{0}^{t_1} \mathbb{E}_s^{t_2}\left(D_s L_{E}(t_1,t_1,t_2) \right) dW^{t_2}_{s}\right)\\
&= L_{E}(0,t_1,t_2) + \mathbb{E}^{t_1}\left( \int_{0}^{t_1} \mathbb{E}_s^{t_2}\left(D_s L_{E}(t_1,t_1,t_2) \right) (\nu(s,t_2)-\nu(s,t_1)) ds \right).
\end{align*}
Now from (\ref{approximation_D_s_x_t}) we have that 
\begin{eqnarray*}
D_s L(t_1,t_1,t_2) &=& \frac{G(t_1,t_2)}{\delta_{t_1,t_2}P_{E}(t_1,t_2)} D_s x_{t_1}\\
&\approx& \frac{G(t_1,t_2)}{\delta_{t_1,t_2}P_{E}(0,t_1,t_2)} \beta(s,t_1, \bar{x}_0(t_1), \bar{y}_s)\bar{DM}(s,t_1)
\end{eqnarray*}
If we use the above approximation, we get
\begin{equation*}
CA(t_0,t_1) = \frac{G(t_1,t_2)}{\delta_{t_1,t_2}P_{E}(0,t_1,t_2)} \beta(s,t_1, \bar{x}_0(t_1), \bar{y}_s) \int_{0}^{t_1} \bar{DM}(s,t_1) (\bar{\nu}(s,t_2,\bar{x}_{0}(t_1))-\bar{\nu}(s,t_1,\bar{x}_{0}(t_1))) ds
\end{equation*}
with $\bar{\nu}(t,t_p, \bar{x}_0(t_a))= \int_{t}^{t_p} \eta(s,\bar{x}_0(t_a),\bar{y}_s) ds$.
